Boosted Regression Trees for Data-Driven Auto-Tuning

### 2011-12-09 paper/draft1.tex
paper

\documentclass{sig-alternate}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage{myAlgorithm}

@@

@@

@@

Following [XXX], the experiments in this paper will focus on filterbank
correlation.  Mathematically, we define filterbank correlation in terms of an
image $x$ and a filterbank $f$.
The image $x$ has $R$ rows, $C$ columns, and $D$ channels (e.g. color
channels) that we call its {\em depth}. We index $x$ like $\mathbf{x}[i,j,d]$
where $0 \leq i < R$, $0 \leq j < C$, and $0 \leq d < D$.
The filterbank $f$ has $F$ filters that are like little images: each has a
height $H$, a width $W$, and $D$ channels.
We will restrict ourselves to what are called {\em valid} correlations, in
which the image is larger in both rows and columns than the filters.
The result of filterbank correlation of $x$ with $f$ is an image-like array
$z$ with $R-H+1$ rows, $C-W+1$ columns, and depth $F$, whose elements are
defined according to Equation~\ref{eq:z}:

@@

Filterbank correlation is used extensively in image and video processing,
where it is often the computational bottleneck.
GPUs are well-suited to computing filterbank correlations, although the
optimal strategy for doing so often depends on the shape of the problem, i.e.
the choices of $R$, $C$, $D$, $F$, and $W$.
The computation of filterbank correlation on GPUs was recently treated by
\citet{pinto+cox:2011gcg}, which serves as the basis for the present work.

@@

@@

\begin{figure}[!ht]
\centerline{
\begin{algorithm}{$\Algo{fbcorr}\big(gX, cF, gZ \big)$}
\Aitem shared $sX \setto$ all channels of region ($\beta$) of $gX$
\Aitem $x, y \setto$ position of this thread in output image
\Aitem \_\_syncthreads()
\Aitem $v[0:N] \setto 0$, for $N=4\times n\_output\_4s$
\Aitem for $d \setto 0$~\To~$D$,
\Aitem ~ for $h \setto 0$~\To~$H / n\_filter\_r$,
\Aitem ~ ~ for $w \setto 0$~\To~$W$,
\Aitem ~ ~ ~ $u \setto sX[x+h, y+w, d]$
\Aitem ~ ~ ~ for $n \setto 0$~\To~$n\_output\_4s - 1$,
\Aitem ~ ~ ~ ~  $v[n] \setto v[n] + cF[n, h, w, d]$
\Aitem for $n \setto 0$~\To~$n\_output\_4s - 1$,
\Aitem ~  gZ[x][y][4n:4n+n] += v[4n:4n+n], (float4)
\end{algorithm}
}
\caption{Kernel pseudo-code for filterbank correlation.
Input $gX$ is a pointer to $x$ in global memory, input $cF$ is a pointer to
$f$ in either constant or texture memory, and output $gZ$ is a pointer to $z$
in global memory.
Each block of threads modifies $4 \times n\_output\_4s$ channels of a
rectangle (called $\beta$ in code listing) within $z$. A grid of blocks covers all rows and columns of $z$.
Multiple calls can be used to apply all filters of a large filterbank $f$ to
$x$.
}
\label{fig:kernel}
\end{figure}

@@

Often however, the entire filterbank often does not fit into the GPU's
constant memory and so $P$ passes are necessary to compute $z$. In such cases,
the number of bytes moved is scaled almost linearly in $P$:
\begin{align}
\mathrm{Bytes} = 4&RCDP \nonumber \\
& + 4KHWD \nonumber \\
& + 8(R-H+1)(C-W+1)KP
% XXX: incrementing requires reading and writing $z$ right? so actually the 4
% above should be an 8? (see related ticket on github for potential
% optimization about this)
\label{eq:bytesP}
\end{align}

@@

\subsection{System Description Space}

 XXXX Leave for future work?

>>>>

\documentclass{sig-alternate}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage{myAlgorithm}
\usepackage{xcolor}

@@

\definecolor{darkred}{rgb}{.5,0,0}

@@

Filterbank correlation is a simple spatial image filtering operation that is
an important subroutine in many image processing applications. It has a relatively
high arithmetic intensity which makes it a natural fit for GPU platforms.
One might expect then, that it would be easy to implement a library providing
this operation -- this hypothetical library could have quite a simple interface
following the spirit of FFTW or a BLAS routine:
\begin{verbatim}
    gpu_fbcorr(
        img_shape, filters_shape,
        img, filters, output)
\end{verbatim}
However, as shown in \citet{pinto+cox:2011gcg} and numerous articles on
related stencil operations XXX, it is challenging to provide an implementation
or even an implementation strategy that provides satisfactory performance
across the range of inputs (shapes, physical layouts) that occur even in
typical usage.

Current state-of-the art is to do empirical auto-tuning.
XXX \vspace{12pt}

Increasingly, machine learning and statistical inference strategies are
finding application in compiler technology.
XXX \vspace{12pt}

This work shows that online auto-tuning can be accelerated by orders of
magnitude by using a regression model built offline as a surrogate for actual
computations on the real hardware.
We use a powerful non-linear regression model (boosted decision trees)
to approximate the empirical relationship between three sets of variables ($\mathcal{A},
\mathcal{B}, \mathcal{C}$) and the observed speed-up over a
reference implementation:
\begin{description}
\item[$\mathcal{A}$:] task description (e.g. shapes, physical layout)
\item[$\mathcal{B}$:] algorithm description (i.e. auto-tuning parameters)
\item[$\mathcal{C}$:] platform description (e.g. micro-benchmarks)
\end{description}
Ideally, \texttt{gpu\_fbcorr} would apply the best
available implementation

\begin{figure}[!ht]
\centerline{
\begin{algorithm}{$\Algo{\color{darkred}{gpu\_fbcorr\_empirical}}\big(shapes, strides \big)$}
\Aitem $a \setto TaskFeatures(shapes, strides)$
\Aitem $c \setto PlatformFeatures()$
\Aitem $b^* \setto \operatorname*{argmin}_{b \in \mathcal{B}} MeasureTime(a, b, c)$
\algoremark{\color{darkred}{slow}}
\Aitem \Return $b^*$
\end{algorithm}
}
\begin{algorithm}{$\Algo{\color{darkred}{gpu\_fbcorr\_predictive}}\big(shapes, strides \big)$}
\Aitem $a \setto TaskFeatures(shapes, strides)$
\Aitem $c \setto PlatformFeatures()$
\Aitem $f \setto LoadTimingModel()$
\Aitem $b^* \setto \operatorname*{argmin}_{b \in \mathcal{B}} f(a, b, c)$
\algoremark{\color{darkred}{fast}}
\Aitem \Return $b^*$
\end{algorithm}
\caption{Pseudo-code template for an auto-tuning implementation of
$\texttt{gpu\_fbcorr}$. Empirical auto-tuning (above) is inevitably slow
because dynamically-generated code must be compiled and run on a number of actual-size inputs.
Predictive auto-tuning (below) can be orders of magnitude faster. We show that it can
also be accurate.
}
\label{fig:argmin}
\end{figure}

machine learning

This paper shows that a kernel with several  boosted

@@

Following \citet{pinto+cox:2011gcg}, the experiments in this paper will focus on filterbank
correlation.  Mathematically, we define filterbank correlation in terms of an
image $x$ and a filterbank $f$.
The image $x$ has $R$ rows, $C$ columns, and $D$ channels (e.g. color
channels) that we call its {\em depth}. We index $x$ like $\mathbf{x}[i,j,d]$
where $0 \leq i < R$, $0 \leq j < C$, and $0 \leq d < D$.
The filterbank $f$ has $F$ filters that are like little images: each has a
height $H$, a width $W$, and $D$ channels.
We will restrict ourselves to what are called {\em valid} correlations, in
which the image is larger in both rows and columns than the filters.
The result of filterbank correlation of $x$ with $f$ is an image-like array
$z$ with $R-H+1$ rows, $C-W+1$ columns, and depth $F$, whose elements are
defined according to Equation~\ref{eq:z}:

@@

Filterbank correlation is used extensively in image and video processing,
where it is often the computational bottleneck.
GPUs are well-suited to computing filterbank correlations, although the
optimal strategy for doing so often depends on the shape of the problem, i.e.
the values of $R$, $C$, $D$, $F$, and $W$.

@@

In terms of artithmetic density, it's approximately
\begin{align}
\mathrm{intensity} & \approx \frac {KDHW} {2(D+K)},
\end{align}
because each $K$ output writes corresponds to approximately $D$ input reads
and an inner product between $KDHW$ elements.

@@

\begin{figure}[!ht]
\centerline{
\begin{algorithm}{$\Algo{thread\_fbcorr}\big(gX, cF, gZ \big)$}
\Aitem shared $sX \setto$ all channels of region ($\beta$) of $gX$
\Aitem $x, y \setto$ position of this thread in output image
\Aitem \_\_syncthreads()
\Aitem $v[0:N] \setto 0$, for $N=4\times n\_output\_4s$
\Aitem for $d \setto 0$~\To~$D$,
\Aitem ~ for $h \setto 0$~\To~$H / n\_filter\_r$,
\Aitem ~ ~ for $w \setto 0$~\To~$W$,
\Aitem ~ ~ ~ $u \setto sX[x+h, y+w, d]$
\Aitem ~ ~ ~ for $n \setto 0$~\To~$n\_output\_4s - 1$,
\Aitem ~ ~ ~ ~  $v[n] \setto v[n] + cF[n, h, w, d]$
\Aitem for $n \setto 0$~\To~$n\_output\_4s - 1$,
\Aitem ~  gZ[x][y][4n:4n+n] += v[4n:4n+n], (float4)
\end{algorithm}
}
\caption{Kernel pseudo-code for filterbank correlation.
Input $gX$ is a pointer to $x$ in global memory, input $cF$ is a pointer to
$f$ in either constant or texture memory, and output $gZ$ is a pointer to $z$
in global memory.
Each block of threads modifies $4 \times n\_output\_4s$ channels of a
rectangle (called $\beta$ in code listing) within $z$. A grid of blocks covers all rows and columns of $z$.
Multiple calls can be used to apply all filters of a large filterbank $f$ to
$x$.
}
\label{fig:kernel}
\end{figure}

@@

When the entire filterbank often does not fit into the GPU's
constant memory and so $$P=\frac{KH}{4 \cdot n\_output\_4s \cdot
n\_filter\_r}$$ passes are necessary to compute $z$. In such cases,
the number of bytes moved is actually more like:
\begin{align}
\mathrm{Bytes} = 4&RCDP \nonumber \\
& + 4KHWD \nonumber \\
& + 8(R-H+1)(C-W+1)KP.
\nonumber
% XXX: incrementing requires reading and writing $z$ right? so actually the 4
% above should be an 8? (see related ticket on github for potential
% optimization about this)
\label{eq:bytesP}
\end{align}
These passes make the I/O requirements increase quadratically in $K$ and $H$.
At the same time, the total number of floating-point operations (Eq.~\ref{eq:flops})
is quadratic in $H$ and $W$. In our experiments, we only considered square
filters so in our setting the total number of flops is proportional to $H^4$.
% XXX WHERE IS ALL THIS GOING?

Best arithmetic density

A bandwidth analysis that takes the blocking structure into account would
reveal that larger $H$ and $W$ require each block to read a larger input
region, and store it in a larger shared memory region, which reduces the
potential for high occupancy.

@@
